Identifying the Tissue of Origin of Extracellular Vesicles Using RNA Expression Signatures

### 2015-09-17 poster/poster.Rnw
switch to diagnostic

    One requirement for remotely sensing the transcriptome is
    identifying the source tissue of a specific \ac{EV}. To identify
    the source tissue of a specific \ac{EV}, we have identified
    tissue-specific miRNA and mRNA markers for multiple organs and
    tissues from existing publicly-available RNAseq transcriptomes by
    calculating a tissue specificity index.

@@

\column{0.25}
  \block{Uterus Specific Genes}{

@@

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Uterus Specific Genes")
      + ylab("Tissue"))
@
}
\captionof{figure}{Expression of uterus specific genes
  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
  highest expression for any of the uterus specific genes ordered by
  their specificity to uterus. The known uterus specific gene PAEP and
  short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp) are highly
  uterus specific.}

@@

  \block{Placenta Specific Genes}{
{\centering
% <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="placenta" & gi.max >= 1000,
%                               list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Placenta specific genes",label="tab:placenta_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @

<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("placenta",
                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
                                                                         gi.max >= 1024
                               ,gene]))
      + xlab("Uterus Specific Genes")
      + ylab("Tissue"))
@

}
\captionof{figure}{Expression of placenta specific genes
  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
  in placenta and all tissues with the second highest expression for
  any placenta-specific gene ordered by their specificity to placenta.
  The previously known placenta-specific PSG genes and CSH genes are
  highly placenta specific and the short RNAs mir-126, mir-622, and
  PAPPA-AS1 are fairly placenta specific.}
}

@@

\block{Conclusions}{
  \begin{itemize}
  \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
    specific to placental tissues, and PAEP is specific to uterus, as
    previously described.
%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
  \item \ac{SVM} are able to accurately identify the tissue of origin
    for uterus and placenta.
  \item Very few sequenced reads are required to accurately identify
    tissues of Placental origin. We are able to detect 100\% of the
    placenta tissues with 80\% specificity with just 10 reads.
  \item We can easily identify exosomes originating from the placenta,
    as even the smallest exosomes can contain 70 RNA molecules
    \cite{Li.ea2014:AnalysisofRNAcontent}.
  \item Non-maternal variants may also be able to distinguish between
    placental reads and maternal reads.
  \end{itemize}
}

>>>>

    One requirement for remotely sensing the transcriptome is
    identifying the source tissue of a specific \ac{EV}. To identify
    the source tissue of a specific \ac{EV}, we have identified miRNA
    and mRNA markers which are diagnostic of multiple organs and
    tissues from existing publicly-available RNAseq transcriptomes by
    calculating a tissue specificity index.

@@

\column{0.25}
  \block{Genes Diagnostic of Uterus}{

@@

<<uterus.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("uterus")
      + xlab("Genes Diagnostic of Uterus ")
      + ylab("Tissue"))
@
}
\captionof{figure}{Expression of genes diagnostic of uterus
  ($\tau_g \ge 0.98$) in uterus and all tissues with the second
  highest expression for any of the uterus-diagnostic genes
  ordered by their tissue specificity. The known uterus-diagnostic
  gene PAEP and short RNAs ZNFX1-AS1\_3 (87bp) and AL356310.1 (94bp)
  are highly diagnostic of uterus.}

@@

  \block{Genes Diagnostic of Placenta}{
{\centering
% <<placenta.specific.markers.table,results="asis",echo=FALSE>>=
% temp <- tissue.specific.genes[tissue=="placenta" & gi.max >= 1000,
%                               list(gene,specificity,gi.max)]
% setorder(temp,-gi.max)
% setnames(temp,c("Gene","Tissue Specificity","Mean Expression"))
% print(xtable(temp,
%              caption="Placenta specific genes",label="tab:placenta_specific_genes"),
%       floating=FALSE,
%       include.rownames=FALSE,
%       )
% rm(temp)
% @

<<placenta.specific.markers.figure,fig.width=8,fig.height=8,echo=FALSE>>=
print(specific.markers.plot("placenta",
                            genes=as.character(tissue.specific.genes[tissue=="placenta" & 
                                                                         gi.max >= 1024
                               ,gene]))
      + xlab("Genes Diagnostic of Placenta")
      + ylab("Tissue"))
@

}
\captionof{figure}{Expression of genes diagnostic of placenta
  ($\tau_g \ge 0.98$) with high expression ($\mathrm{FPKM} \ge 1024$)
  in placenta and all tissues with the second highest expression for
  any placenta-diagnostic gene ordered by their specificity to
  placenta. The previously known placenta-diagnostic PSG genes and
  CSH genes are nearly unique to placenta and the short RNAs mir-126,
  mir-622, and PAPPA-AS1 are less so.}
}

@@

\block{Conclusions}{
  \begin{itemize}
  \item Pregnancy-specific glycoproteins (PSG1, 2, 3 et al.) are
    diagnostic of placental tissues, and PAEP is diagnostic of
    uterus, as previously described.
%   \item Numerous Long-noncoding RNA and smaller RNAs are specific to uterus and placenta
  \item \ac{SVM} are able to accurately identify the tissue of origin
    for uterus and placenta.
  \item Very few sequenced reads are required to accurately identify
    tissues of Placental origin. We are able to detect 100\% of the
    placenta tissues with 80\% specificity with just 10 reads.
  \item We can easily identify exosomes originating from the placenta,
    as even the smallest exosomes can contain 70 RNA molecules
    \cite{Li.ea2014:AnalysisofRNAcontent}.
  \item Non-maternal variants may also be able to distinguish between
    placental reads and maternal reads.
  \end{itemize}
}
